---
title: "Length_Weight_practice"
author: "Jillian Burns"
date: "11/4/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r echo=FALSE, warning=FALSE, message=FALSE}
library(reshape2)
library(dplyr)
library(tidyr)
library(ggplot2)
require(scales)
```


# Introduction

In order to calculate gut fullness for my fish, I need to calcuate fish weights from a length-weight relationship because the scale I used was not appropriately precise. SS from CDFW provided me length-weight data for larval longfin smelt collected from the IEP Smelt Larval Survey to get this relationship. Find more information about the [Smelt Larval Survey here](https://www.dfg.ca.gov/delta/projects.asp?ProjectID=SLS). 

```{r echo=FALSE}
# read in dataset
a <- read.csv("CDFW LONSME L-W_5Aug2019.csv", stringsAsFactors = F)
head(a)

# format date in order to filter
a$SurveyDate <- as.Date(a$SurveyDate, "%d-%b-%y")
a$Month <- format(as.Date(a$SurveyDate), "%b")

# filter for only relevant fish lengths and dates for my dataset
b <- a %>% #n=304 (b dataframe)
  filter(ForkLength %in% (4:28) & Month %in% c("Jan", "Feb", "Mar", "Apr"))
b$Month <- factor(b$Month, levels = c("Jan", "Feb", "Mar", "Apr"))
```


# Analysis

First, I want to see the raw data.  
```{r}
hist(b$ForkLength)

```
```{r}
plot(b$ForkLength, b$TotalBodyWeight)
```


# Results

```{r}
#Start plotting - change data here depending on what length range you want
(p <- ggplot(b, aes(x = Month, y = ForkLength)))
#Plot raw values
(pp <- p + geom_boxplot()) +
  theme_bw()
```

